Windowed least-squares model reduction for dynamical systems

### 2019-10-18 main.tex
latest

Model reduction for linear-time-invariant systems (and other well structured
dynamical systems) is quite mature
\cite{wilcox_benner_rev,moore,roberts,GugercinIRKA}, as system-theoretic properties (e.g., controllability,
observability, asymptotic stability, $\mathcal H_2$-optimality) can be readily
quantified and accounted for; this often results in
reduced-order models that inherit such important properties.  The primary
challenge in
developing reduced-order models for general nonlinear dynamical systems is
that such properties are difficult to assess quantitatively. As a result, it
is difficult to develop reduced-order models that preserve important
dynamical-system properties, which often results in methods that yield
trajectories that are
inaccurate, unstable, or violate physical properties.  To address this, researchers have pursued several directions that
aim to imbue reduced-order models for nonlinear
dynamical systems with properties that can improve robustness and accuracy.
These
efforts include residual-minimization approaches that equip the ROM solution with a notion of optimality~\cite{carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis,l1};
space--time approaches that leads to error bounds that grow slowly in time~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st};
``energy-based'' inner
products that ensure non-increasing entropy in the ROM
solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014,chan2019entropy};
basis-adaptation methods that improve the ROM's accuracy \textit{a
posteriori}~\cite{carlberg_hadaptation,adeim_peherstorfer,etter2019online}, stabilizing subspace rotations that account for truncated modes \textit{a priori}~\cite{basis_rotation}, structure-preserving
methods that enforce conservation \cite{carlberg_conservative_rom} or
(port-)Hamiltonian/Lagrangian structure
\cite{LALL2003304,carlberg2012spd,structurePreserveBeattie,chaturantabut2016structure,farhat2014dimensional} in the ROM; and subgrid-scale
modeling methods that aim to improve accuracy by addressing the closure
problem~\cite{san_iliescu_geostrophic,iliescu_pod_eddyviscosity,iliescu_vms_pod_ns,Bergmann_pod_vms,iliescu_ciazzo_residual_rom,Wang_ROM_thesis,wentland_apg,Wang:269133,San2018}.
We note that these techniques are often not mutually exclusive.
%Residual-minimizing projection methods are the most relevant to the current work and comprise the focus of the following review.
Residual-minimizing and space--time approaches are the most relevant classes of
methods for the current work and comprise the focus of the following review.

Residual-minimization methods in model reduction compute the solution within a
low-dimensional trial subspace that minimizes the full-order-model residual.
Researchers have developed such residual-minimizing model-reduction methods
for both static systems (i.e., systems without time-dependence)
\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
and dynamical
systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
In the latter category,
Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
formulated the residual minimization problem for dynamical systems by
sequentially minimizing the \textit{time-discrete} full-order-model residual
(i.e., the residual arising after applying time discretization) at each time
instance on the time-discretization grid. This formulation is often referred
to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
formulation and examined its connections with Galerkin projection. Critically,
this work demonstrated that (1) under certain conditions, LSPG can be cast as a 
Petrov--Galerkin projection applied to the time-continuous
full-order-model residual, and (2) LSPG and Galerkin projection are equivalent
in the limit as the time step goes to zero (i.e., Galerkin projection 
minimizes the time-instantaneous full-order-model residual). Numerous
numerical experiments have demonstrated that LSPG often yields more
accurate and stable solutions than Galerkin
~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
The common intuitive explanation for this improved performance is that, by
minimizing the full-order-model residual over a finite time window (rather
than time instantaneously), LSPG computes solutions that are more
accurate over a larger part of the trajectory as compared to Galerkin.

@@

Space--time residual minimization approaches minimize the FOM residual over all of space and time and, as a result, yield models that are equipped with a notion of space--time optimality and \textit{a priori} error bounds that grow more slowly
in time. Further, space--time approaches reduce both the spatial and temporal dimensions of the full-order-model, and thus promise cost savings over spatial-projection-only approaches.
% and the resulting \textit{trajectory} exhibits a notion of
%optimality; this implies that the solution exhibits a notion of optimality
%over the entire time domain. 
%\EP{Come back to this, this also has to do with residual minimization}
However, space--time techniques also suffer from several
limitations. First, the computational cost of solving the algebraic system arising from space--time
approaches scales cubically with the number of space--time degrees of freedom;
in contrast, the computational cost incurred by standard
spatial-projection-based ROMs is linear in the number of temporal degrees of
freedom, as the attendant solvers can leverage the lower-block triangular
structure of the system arising from the sequential nature of time
evolution. As a result, solving the algebraic systems arising from space--time
projection is generally intractable without applying hyper-reduction in time
\cite{choi_stlspg,constantine_strom}. Second, space--time residual
minimization precludes future
state prediction, as these methods employ space--time basis vectors defined over the
entire time domain of interest, which must have been included in the training
simulations.

@@

\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.1cm 0cm 1cm 0cm},clip,width=0.99\textwidth]{diagram.pdf} 
\caption{Relationship diagram for the \methodAcronym\ approach for model
	reduction. \KTC{Also, can we make this more specific? In particular, can
	we write specific equations (e.g., ``$W = T$ in Eq.~(10)'') for the limiting
	cases?} \EP{I tried, it becomes very crowded. I think with the description of trial spaces its pretty good now}} 
\label{fig:flowchart} 
\end{centering} 
\end{figure}

 In summary, specific contributions of this work include:
\begin{enumerate}
%\item The development of a time-continuous residual minimization framework for reduced-order models.
\item The \methodNameLower\ (\methodAcronym) approach for dynamical-system
	model reduction. The approach sequentially minimizes the time-continuous
		full-order model residual over arbitrary time windows.
%\item Support of two space--time trial subspaces: one that reduces the spatial
%	dimension of the FOM only, and one that reduces both the spatial and
%		temporal dimensions of the FOM. 
\item Support of two space--time trial subspaces: one that associates with spatial dimension reduction and 
    one that associates with spatial and temporal dimension reduction.
The former case is of particular
		interest in the \methodAcronym\  contex, as the stationary conditions are
		derived via the Euler--Lagrange equations and comprise a coupled two-point
		Hamiltonian boundary value problem containing a forward and backward
		system. The forward system, which is forced by an auxiliary costate,
		evolves the (spatial) generalized coordinates of the ROM in time. The
		backward system, which is forced by the time-continuous FOM residual
		evaluated about the ROM state, governs the dynamics of the costate. 
	\item Derivation of two solution techniques:
		\textit{discretize-then-optimize} (i.e., direct) and
		\textit{optimize-then-discretize} (i.e., indirect). 
	\item Remarks and derivations of conditions under which the
		\methodAcronym\ approach recovers Galerkin, LSPG, and space--time LSPG.
\item Error analysis of the \methodAcronym\ approach using
	trial subspaces associated with spatial dimension reduction.
	This analysis demonstrates that, over a given window, the \methodAcronymROMs\ error 
		is bounded \textit{a priori} by a combination of the
		error at the start of the window and the integrated FOM ODE residual
		evaluated at the FOM state projected onto the trial subspace. 
\EP{Updated this, I don't show what you said}%As a result,
%		the error bound exhibits exponential growth in the number of time windows
%		(not total time instances), yielding smaller error-bound growth in thime than
%		Galerkin or LSPG projection. \KTC{Check this}
\item Numerical experiments for 
	trial subspaces associated with spatial dimenion reduction, which demonstrate two key findings:
\begin{itemize}
\item Minimizing the residual over a larger time window leads to more stable solutions 
with lower space--time residuals norms.
\item Minimizing the residual over a larger time window does not necessarily
	lead to a more accurate trajectory (as measured in the space--time
		$\elltwo$-norm of the solution). Instead, minimizing the residual over an
		intermediate-sized time window leads to the smallest trajectory error.
\end{itemize}
\end{enumerate}

>>>>

Model reduction for linear-time-invariant systems (and other well structured
dynamical systems) is quite mature
\cite{wilcox_benner_rev,moore,roberts,GugercinIRKA}, as system-theoretic properties (e.g., controllability,
observability, asymptotic stability, $\mathcal H_2$-optimality) can be readily
quantified and accounted for; this often results in
reduced-order models that inherit such important properties.  The primary
challenge in
developing reduced-order models for general nonlinear dynamical systems is
that such properties are difficult to assess quantitatively. As a result, it
is challenging to develop reduced-order models that preserve important
dynamical-system properties, which often results in methods that yield
trajectories that are
inaccurate, unstable, or violate physical properties.  To address this, researchers have pursued several directions that
aim to imbue reduced-order models for nonlinear
dynamical systems with properties that can improve robustness and accuracy.
These
efforts include residual-minimization approaches that equip the ROM solution with a notion of optimality~\cite{carlberg_lspg,carlberg_gnat,legresley_1,legresley_2,legresley_3,bui_resmin_steady,bui_unsteady,rovas_thesis,carlberg_thesis,bui_thesis,l1};
space--time approaches that lead to error bounds that grow slowly in time~\cite{choi_stlspg,constantine_strom,URBAN2012203,Yano2014ASC,benner_st};
``energy-based'' inner
products that ensure non-increasing entropy in the ROM
solution~\cite{rowley_pod_energyproj,Kalashnikova_sand2014,chan2019entropy};
basis-adaptation methods that improve the ROM's accuracy \textit{a
posteriori}~\cite{carlberg_hadaptation,adeim_peherstorfer,etter2019online}, stabilizing subspace rotations that account for truncated modes \textit{a priori}~\cite{basis_rotation}, structure-preserving
methods that enforce conservation \cite{carlberg_conservative_rom} or
(port-)Hamiltonian/Lagrangian structure
\cite{LALL2003304,carlberg2012spd,structurePreserveBeattie,chaturantabut2016structure,farhat2014dimensional} in the ROM; and subgrid-scale
modeling methods that aim to improve accuracy by addressing the closure
problem~\cite{san_iliescu_geostrophic,iliescu_pod_eddyviscosity,iliescu_vms_pod_ns,Bergmann_pod_vms,iliescu_ciazzo_residual_rom,Wang_ROM_thesis,wentland_apg,Wang:269133,San2018}.
We note that these techniques are often not mutually exclusive.
%Residual-minimizing projection methods are the most relevant to the current work and comprise the focus of the following review.
Residual-minimizing and space--time approaches are the most relevant classes of
methods for the current work and comprise the focus of the following review.

Residual-minimization methods in model reduction compute the solution within a
low-dimensional trial subspace that minimizes the full-order-model residual.\footnote{While we focus our review on residual-minimization approaches in the context of model reduction, we note that these approaches are intimately related to least-squares finite element methods,~\cite{bochev_leastsquares,bochev_lsfem_book} which predate ROM residual-minimization approaches.}
Researchers have developed such residual-minimizing model-reduction methods
for both static systems (i.e., systems without time-dependence)
\cite{legresley_1,legresley_2,legresley_3,bui_resmin_steady,rovas_thesis,carlberg_thesis,bui_thesis}
and dynamical
systems~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg,carlberg_lspg_v_galerkin}.
In the latter category,
Refs.~\cite{bui_thesis,bui_unsteady,carlberg_thesis,carlberg_gnat,carlberg_lspg}
formulated the residual minimization problem for dynamical systems by
sequentially minimizing the \textit{time-discrete} full-order-model residual
(i.e., the residual arising after applying time discretization) at each time
instance on the time-discretization grid. This formulation is often referred
to as the \textit{least--squares Petrov--Galerkin} (LSPG) method.
Ref.~\cite{carlberg_lspg_v_galerkin} performed detailed analyses of this
formulation and examined its connections with Galerkin projection. Critically,
this work demonstrated that (1) under certain conditions, LSPG can be cast as a 
Petrov--Galerkin projection applied to the time-continuous
full-order-model residual, and (2) LSPG and Galerkin projection are equivalent
in the limit as the time step goes to zero (i.e., Galerkin projection 
minimizes the time-instantaneous full-order-model residual). Numerous
numerical experiments have demonstrated that LSPG often yields more
accurate and stable solutions than Galerkin~\cite{bui_thesis,carlberg_lspg_v_galerkin,carlberg_gnat,carlberg_thesis,parish_apg}.
The common intuitive explanation for this improved performance is that, by
minimizing the full-order-model residual over a finite time window (rather
than time instantaneously), LSPG computes solutions that are more
accurate over a larger part of the trajectory as compared to Galerkin.

@@

Space--time residual minimization approaches minimize the FOM residual over all of space and time and, as a result, yield models that are equipped with a notion of space--time optimality and \textit{a priori} error bounds that grow more slowly
in time. Further, space--time approaches reduce both the spatial and temporal dimensions of the full-order model, and thus promise cost savings over spatial-projection-only approaches.
% and the resulting \textit{trajectory} exhibits a notion of
%optimality; this implies that the solution exhibits a notion of optimality
%over the entire time domain. 
%\EP{Come back to this, this also has to do with residual minimization}
However, space--time techniques also suffer from several
limitations. First, the computational cost of solving the algebraic system arising from space--time
approaches scales cubically with the number of space--time degrees of freedom;
in contrast, the computational cost incurred by standard
spatial-projection-based ROMs is linear in the number of temporal degrees of
freedom, as the attendant solvers can leverage the lower-block triangular
structure of the system arising from the sequential nature of time
evolution. As a result, solving the algebraic systems arising from space--time
projection is generally intractable without applying hyper-reduction in time
\cite{choi_stlspg,constantine_strom}. Second, space--time residual
minimization precludes future
state prediction, as these methods employ space--time basis vectors defined over the
entire time domain of interest, which must have been included in the training
simulations.

@@

\begin{figure} 
\begin{centering} 
\includegraphics[trim={0.1cm 0cm 1cm 0cm},clip,width=0.99\textwidth]{diagram.pdf} 
\caption{Relationship diagram for the \methodAcronym\ approach for model
	reduction.} 
\label{fig:flowchart} 
\end{centering} 
\end{figure}

 In summary, specific contributions of this work include:
\begin{enumerate}
%\item The development of a time-continuous residual minimization framework for reduced-order models.
\item The \methodNameLower\ (\methodAcronym) approach for dynamical-system
	model reduction. The approach sequentially minimizes the time-continuous
		full-order-model residual over arbitrary time windows.
%\item Support of two space--time trial subspaces: one that reduces the spatial
%	dimension of the FOM only, and one that reduces both the spatial and
%		temporal dimensions of the FOM. 
\item Support of two space--time trial subspaces: one that associates with spatial dimension reduction and 
    one that associates with spatial and temporal dimension reduction.
The former case is of particular
		interest in the \methodAcronym\  contex, as the stationary conditions are
		derived via the Euler--Lagrange equations and comprise a coupled two-point
		Hamiltonian boundary value problem containing a forward and backward
		system. The forward system, which is forced by an auxiliary costate,
		evolves the (spatial) generalized coordinates of the ROM in time. The
		backward system, which is forced by the time-continuous FOM residual
		evaluated about the ROM state, governs the dynamics of the costate. 
	\item Derivation of two solution techniques:
		\textit{discretize then optimize} (i.e., direct) and
		\textit{optimize then discretize} (i.e., indirect). 
	\item Remarks and derivations of conditions under which the
		\methodAcronym\ approach recovers Galerkin, LSPG, and space--time LSPG.
\item Error analysis of the \methodAcronym\ approach using
	trial subspaces associated with spatial dimension reduction.
	This analysis demonstrates that, over a given window, the \methodAcronymROMs\ error 
		is bounded \textit{a priori} by a combination of the
		error at the start of the window and the integrated FOM ODE residual
		evaluated at the FOM state projected onto the trial subspace. 
\EP{Updated this, I don't show what you said}%As a result,
%		the error bound exhibits exponential growth in the number of time windows
%		(not total time instances), yielding smaller error-bound growth in thime than
%		Galerkin or LSPG projection. \KTC{Check this}
\item Numerical experiments for 
	trial subspaces associated with spatial dimenion reduction, which demonstrate two key findings:
\begin{itemize}
\item Minimizing the residual over a larger time window leads to more stable solutions 
with lower space--time residuals norms.
\item Minimizing the residual over a larger time window does not necessarily
	lead to a more accurate trajectory (as measured in the space--time
		$\elltwo$-norm of the solution). Instead, minimizing the residual over an
		intermediate-sized time window leads to the smallest trajectory error.
\end{itemize}
\end{enumerate}

### 2019-10-18 section3/space_time.tex
latest

%We now briefly consider \spaceTimeAcronym\ trial subspaces. 
The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
the FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$
with $\approxstateArgnt{n} \in \stspaceSTArg{n}$, where
\begin{equation}\label{eq:st_sttrialspace}
 \stspaceSTArg{n} \defeq
 \Range{\stbasisArgnt{n}} + \stateInterceptSTArg{n}\otimes \onesFunctionArg{n} \subseteq \RR{\fomdim} \otimes \timeSpaceArg{n}.
\end{equation}
\EP{Is it wrong to say $\stbasisArg{n}{\cdot} \in \RR{\fomdim \times \stdimArg{n}}$ with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy) \ldots$? I don't particularly like
$\stbasisArgnt{n} \in \RR{\fomdim \times \stdimArg{n}}\otimes\timeSpaceArg{n}$; as the reader I think that $\stbasisArgnt{n}$ can take on all temporal functions.}
Here $\stbasisArgnt{n} \in \RR{\fomdim \times
\stdimArg{n}}\otimes\timeSpaceArg{n}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{\timeStartArg{n}} = \bz$ is the space--time trial basis matrix {function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$, $n=1,\ldots,\nslabs$ provides the affine transformation. 
To enforce the initial condition and ensure solution continuity across time windows, we set $\stateInterceptSTArg{1} = \stateFOMIC$ and
$\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for
$n=2,\ldots,\nslabs$.
At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as
 \begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},
\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window. 
\begin{comment}
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptSTArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
\KTC{Remove all commas before equations unless it makes sense gramatically.
Add the $n$ superscript to $x_{ref}$}
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
%where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

>>>>

%We now briefly consider \spaceTimeAcronym\ trial subspaces. 
The \spaceTimeAcronym\ trial subspace over the $n$th time window approximates
the FOM ODE solution trajectory $\stateFOM\in\RR{N}\otimes\timeSpace$
with $\approxstateArgnt{n} \in \stspaceSTArg{n}$, where
\begin{equation}\label{eq:st_sttrialspace}
 \stspaceSTArg{n} \defeq
 \Range{\stbasisArgnt{n}} + \stateInterceptSTArg{n}\otimes \onesFunctionArg{n} \subseteq \RR{\fomdim} \otimes \timeSpaceArg{n}.
\end{equation}
Here $\stbasisArgnt{n} \in \RR{\fomdim \times
\stdimArg{n}}\otimes\timeSpaceArg{n}$, $n=1,\ldots,\nslabs$, with $\stbasisArgnt{n}: \timeDummy \mapsto \stbasisArgnt{n}(\timeDummy)$ and $\stbasisArg{n}{\timeStartArg{n}} = \bz$ is the space--time trial basis matrix {function} and $\stateInterceptSTArg{n} \in \RR{\fomdim}$, $n=1,\ldots,\nslabs$ provides the affine transformation. 
To enforce the initial condition and ensure solution continuity across time windows, we set $\stateInterceptSTArg{1} = \stateFOMIC$ and
$\stateInterceptSTArg{n} = \approxstateArg{n-1}{\timeEndArg{n-1}}$ for
$n=2,\ldots,\nslabs$.
At any time instance $t \in [\timeStartArg{n},\timeEndArg{n}]$, the \spaceTimeAcronym\ trial subspace approximates the FOM ODE solution as
 \begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{n},
\end{equation}
where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window. 
\begin{comment}
each window, we introduce the space--time trial subspace
$$ \stateArgnt{n} \in \stspaceSTArg{n} \subseteq \RR{N} \otimes \timeSpaceArg{n},$$
along with the space--time trial basis matrix \textit{function}, 
\begin{align*}
\stbasisArgnt{n} &: t \mapsto \stbasis(t) \\
 &: [\timeStartArg{n},\timeEndArg{n}] \rightarrow  \RR{N \times \stdimArg{n}}  ,%\spaceTrialSpace , \\
\end{align*}
\KTC{shouldn't be $\equiv$. Look earlier for this. the right hand side should
be on the left with $\defeq$.} with $\Range{\stbasisArgnt{n}} + \stateInterceptSTArg{} \equiv \stspaceSTArg{n} $ and where $\stdimArg{n}$ is the number of space--time generalized coordinates over the $n$th window. 
For simplicity we assume the reference state to be equivalent for each time window, although no such requirement is necessary. 
\KTC{Remove all commas before equations unless it makes sense gramatically.
Add the $n$ superscript to $x_{ref}$}
The state over each window is approximated by,
\begin{equation}\label{eq:stapprox}
 \stateFOMArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstateArg{n} + \stateInterceptSTArg{},
\end{equation}
%and the state is then approximated by,
%\begin{equation}\label{eq:stapprox}
% \stateArg{n}{t} \approx \approxstateArg{n}{t}  = \stbasisArg{n}{t} \stgenstate + \stateInterceptArg{n},
%\end{equation}
%where $\stgenstateArg{n} \in \RR{\stdimArg{n}}$ are the space--time generalized coordinates over the $n$th window.

### 2019-10-18 section4/numerical_methods.tex
latest

\section{Numerical solution techniques}\label{sec:numerical_techniques}
We now consider numerical-solution techniques for the \methodAcronym\ \approachKwd.
We focus primarily on 
direct (i.e., discretize-then-optimize) and indirect (i.e., optimize-then-discretize) methods for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
%It is noted that, as \methodAcronym\ with \spatialAcronym\ trial subspaces 
%Given the formulation of the \methodAcronym\ method presented in Section
%\ref{sec:tclspg}, we now consider numerical-solution techniques for the method. 
%Due to the commonalities that the \methodAcronym\ method exhibits with
%optimal-control problems, \KTC{is this only for S-reduction subspaces? If so,
%state this.} the derivation of these techniques can leverage a large existing
%body of work. In this section, we outline a variety of such techniques. 
\subsection{\spatialAcronym: direct and indirect methods}
\methodAcronym\ with \spatialAcronym\ trial subspaces can be viewed as an optimal-control problem.
Numerical solution techniques for
optimal-control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}. Rather than working with the first-order optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}, direct methods ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the system \KTC{What does it mean to discretize the
system?} and objective functional and ``transcribing"
the infinite-dimensional problem into a finite-dimensional 
problem. Direct approaches thus
``discretize then optimize''. \KTC{Note: only use hyphens when
discretize-then-optimize is used as an adjective.}
In contrast, indirect methods compute solutions to the Euler--Lagrange
equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the
first-order optimality conditions). \KTC{Note: they are not first-order
optimality conditions of the objective functional only; they are first-order
optimality conditions of the optimization problem} Thus, indirect methods
``optimize then discretize," and solve the optimization problem
``indirectly.'' A variety of both discretization and solution techniques are possible
for both direct and indirect methods. Collocation methods,
finite-element methods, spectral methods, and shooting methods are all examples of
possible solution techniques.

>>>>

\section{Numerical solution techniques}\label{sec:numerical_techniques}
We now consider numerical-solution techniques for the \methodAcronym\ \approachKwd.
We focus primarily on 
direct (i.e., discretize-then-optimize) and indirect (i.e., optimize-then-discretize) methods for \spatialAcronym\ trial subspaces, 
and then briefly outline direct and indirect methods for \spaceTimeAcronym\ trial subspaces. 
%It is noted that, as \methodAcronym\ with \spatialAcronym\ trial subspaces 
%Given the formulation of the \methodAcronym\ method presented in Section
%\ref{sec:tclspg}, we now consider numerical-solution techniques for the method. 
%Due to the commonalities that the \methodAcronym\ method exhibits with
%optimal-control problems, \KTC{is this only for S-reduction subspaces? If so,
%state this.} the derivation of these techniques can leverage a large existing
%body of work. In this section, we outline a variety of such techniques. 
\subsection{\spatialAcronym: direct and indirect methods}
\methodAcronym\ with \spatialAcronym\ trial subspaces can be viewed as an optimal-control problem.
Numerical solution techniques for
optimal-control problems can be classified as either
\textit{direct} or \textit{indirect}
methods~\cite{conway_optimalcontrolreview}. Rather than working with the first-order optimality conditions~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs}, direct methods ``directly" solve the optimization problem~\eqref{eq:tclsrm} by first
numerically discretizing the objective functional and ``transcribing"
the infinite-dimensional problem into a finite-dimensional 
problem. Direct approaches thus
``discretize then optimize''. 
In contrast, indirect methods compute solutions to the Euler--Lagrange
equations~\eqref{eq:lspg_continuous}--\eqref{eq:lspg_bcs} (which comprise the
first-order optimality conditions). Thus, indirect methods
``optimize then discretize," and solve the optimization problem
``indirectly.'' A variety of both discretization and solution techniques are possible
for both direct and indirect methods. Collocation methods,
finite-element methods, spectral methods, and shooting methods are all examples of
possible solution techniques.
